The Impact of News Sentiment on MSFT

### 2023-12-02 MSFT.Rmd
ARMA prediction

---
title: "The Impact of ChatGPT news on MSFT"
author: "Ansam Zedan, Daniel Wullschleger"
date: "`r Sys.Date()`"
output: pdf_document
---

@@

```{r include=FALSE}
if (!require("vars")) install.packages("vars", dependencies=TRUE)
if (!require("tseries")) install.packages("tseries", dependencies=TRUE)
if (!require("quantmod")) install.packages("quantmod", dependencies=TRUE)
if (!require("PerformanceAnalytics")) install.packages("PerformanceAnalytics", dependencies=TRUE)
if (!require("forecast")) install.packages("forecast", dependencies=TRUE)
if (!require("rugarch")) install.packages("rugarch", dependencies=TRUE)

library(PerformanceAnalytics)
library(forecast)
library(rugarch)
library(vars)
library(tseries)
library(quantmod)
library(tinytex)
```
# Insert introduction here

@@

#ChatGPT <- read.csv("ChatGPT.csv")

plot(MSFT)
```
We see a clear trend, which is an indicator for non-stationarity, so we do an Augmented Dickey-Fuller (ADF) Test to see whether the data is already stationary.

@@

```{r echo=FALSE, include=TRUE}
MSFT_returns <- na.omit(diff(log(MSFT)))
colnames(MSFT_returns)<-"MSFT_Returns"
adf.test(MSFT_returns)

plot(MSFT_returns)
```
Now, that we deal with stationary data we can begin with our analysis.

First we include the ChatGPT News Time Series to then create a Vector Autoregression model as we want to find out, whether ChatGPT News are a causal driver for MSFT price returns.

@@

```{r echo=FALSE, include=TRUE}
#MSFT_VAR <- VAR(cbind(MSFT_returns, ChatGPT), ic="AIC", lag.max = 24)
#coeftest(MSFT_VAR)
```

As not all coeffcients are signifcant, we may do a causality test. This way, we can find out, if ChatGPT News are in fact a causal driver for MSFT prices

@@

Lastly, it would be interesting to find out how ChatGPT News actually influence the MSFT price returns. So that we may estimate long  such news do have effect on the price.

@@

@@

From class we learned, that in Finance is it common to stick to a GARCH(1,1) model, so we just need to estimate the ARMA model orders. For this, we let auto.arima find the best fitting orders

@@

AS a last step, we may forecast the returns and volatility values.
Here, ARMA will estimate the returns, while GARCH estimates the volatilities

>>>>

---
title: "The Impact of News Sentiment on MSFT"
author: "Ansam Zedan, Daniel Wullschleger"
date: "`r Sys.Date()`"
output: pdf_document
---

@@

```{r include=FALSE}
if (!require("vars")) install.packages("vars", dependencies=TRUE)
if (!require("tseries")) install.packages("tseries", dependencies=TRUE)
if (!require("quantmod")) install.packages("quantmod", dependencies=TRUE)
if (!require("PerformanceAnalytics")) install.packages("PerformanceAnalytics", dependencies=TRUE)
if (!require("forecast")) install.packages("forecast", dependencies=TRUE)
if (!require("rugarch")) install.packages("rugarch", dependencies=TRUE)
if (!require("tinytex")) install.packages("tinytex", dependencies=TRUE)
if (!require("zoo")) install.packages("zoo", dependencies=TRUE)
if (!require("ggplot2")) install.packages("ggplot2", dependencies=TRUE)

library(PerformanceAnalytics)
library(forecast)
library(rugarch)
library(vars)
library(tseries)
library(quantmod)
library(tinytex)
library(zoo)
library(ggpl)
```
# Insert introduction here

@@

#ChatGPT <- read.csv("ChatGPT.csv")
plot(MSFT)
```
We see a clear trend, which is an indicator for non-stationarity, so we do an Augmented Dickey-Fuller (ADF) Test to see whether the data is already stationary.

@@

In the same step, we divide the dataset into a Train/Test Split, whereas November 2023 will serve as a test set for later predicitions

```{r echo=FALSE, include=TRUE}
MSFT_returns <- na.omit(diff(log(MSFT)))
colnames(MSFT_returns)<-"MSFT_Returns"

MSFT_returns_test <- window(MSFT_returns, start="2023-10-01")
MSFT_returns_train <- window(MSFT_returns, end="2023-09-30")

plot(MSFT_returns)
adf.test(MSFT_returns)
```
Now, that we deal with stationary data we can begin with our analysis.

First we include the News Time Series to then create a Vector Autoregression model as we want to find out, whether News are a causal driver for MSFT price returns.

@@

```{r echo=FALSE, include=TRUE}
#MSFT_VAR <- VAR(cbind(MSFT_returns_train, ChatGPT), ic="AIC", lag.max = 24)
#coeftest(MSFT_VAR)
```

As not all coeffcients are signifcant, we may do a causality test. This way, we can find out, if  News are in fact a causal driver for MSFT prices

@@

Lastly, it would be interesting to find out how  News actually influence the MSFT price returns. So that we may estimate long  such news do have effect on the price.

@@

As a benchmark for our VAR Model that features News sentiment as a predictor, we use an ARMA model that is purely based on the price returns of MSFT. This way, we can forecast values of November 2023 and compare the predictions to the actual MSFT price returns to measure its accuracy.

MSFT_prediction <- predict

```{r echo=FALSE, include=TRUE}
MSFT_benchmark_ARMA <- auto.arima(MSFT_returns_train)
MSFT_benchmark_pred <- forecast(MSFT_benchmark_ARMA, level=95, h=43)

plot(MSFT_benchmark_pred)

@@

From class we learned, that in Finance it is common to stick to a GARCH(1,1) model, so we just need to estimate the ARMA model orders. For this, we let auto.arima find the best fitting orders

@@

As a last step, we may forecast the returns and volatility values.
Here, ARMA will estimate the returns, while GARCH estimates the volatilities
